---
title: "reproducibke research"
author: "yakana yakana"
date: "April 4, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```


we are going to load the data using the read.csv package

Data set is gotten from the activity monitoring data 

```{r}
echo=T
cls = c("integer", "character", "integer")
df <- read.csv("activity.csv", head=TRUE, colClasses=cls, na.strings="NA")
head(df)
```
Next step is to process/transform the data set for later analysis.
Specifically, the type of date column is corrected, we also get rid of rows containing missing values and save the subset to a new data frame "df_cor". The original data frame is kept for later data imputation.

```{r}
echo=T
df$date <- as.Date(df$date)
df_cor <- subset(df, !is.na(df$steps))
```

## What is mean total number of steps taken per day?

We draw a histogram of the daily number of steps taking, and this shows a distribution of these totals.

```{r}
echo=T
dailysum <- tapply(df_cor$steps, df_cor$date, sum, na.rm=TRUE, simplify=T)
dailysum <- dailysum[!is.na(dailysum)]

hist(x=dailysum,
     col="blue",
     breaks=30,
     xlab="Daily total steps",
     ylab="Frequency",
     main="The distribution of daily total (missing data ignored)")
```

We then report the mean and median of the total number of steps taking per day.

### mean

```{r}
echo=T
mean(dailysum)
```

### median

```{r}
echo=T
median(dailysum)
```

## What is the average daily activity pattern?

In order to the average daily pattern we create a time series plot of the 5 minute interval (x-axis) and the average number of steps taking, averaged across all the days (Y-axis)

```{r}
echo=T
int_avg <- tapply(df_cor$steps, df_cor$interval, mean, na.rm=TRUE, simplify=T)
df_ia <- data.frame(interval=as.integer(names(int_avg)), avg=int_avg)

with(df_ia,
     plot(interval,
          avg,
          type="l",
          xlab="5-minute intervals",
          ylab="average steps in the interval across all days"))
```

Next is to check which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps:

```{r}
echo=T
max_steps <- max(df_ia$avg)
df_ia[df_ia$avg == max_steps, ]
```

## Imputing missing values
We use a simple strategy for filling in all of the missing values in the dataset. If a 5-minute interval has missing value, we use the mean for that 5-minute interval.

We create a new data frame df_impute that is equal to the original dataset but with the missing data filled in (using mean for that interval for imputation):

```{r}
echo = T
df_impute <- df
ndx <- is.na(df_impute$steps)
int_avg <- tapply(df_cor$steps, df_cor$interval, mean, na.rm=TRUE, simplify=T)
df_impute$steps[ndx] <- int_avg[as.character(df_impute$interval[ndx])]
```

Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day.

```{r}
echo = T
new_dailysum <- tapply(df_impute$steps, df_impute$date, sum, na.rm=TRUE, simplify=T)

hist(x=new_dailysum,
     col="green",
     breaks=20,
     xlab="daily steps",
     ylab="frequency",
     main="The distribution of daily total (with missing data imputed)")
```

## Getting mean and median total number of steps taken per day

```{r}
echo = T
mean(new_dailysum)
median(new_dailysum)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
echo =T
# helper function to decide if a day is a week day or not
is_weekday <- function(d) {
    wd <- weekdays(d)
    ifelse (wd == "Saturday" | wd == "Sunday", "weekend", "weekday")
}

wx <- sapply(df_impute$date, is_weekday)
df_impute$wk <- as.factor(wx)
head(df_impute)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
echo =T
wk_df <- aggregate(steps ~ wk+interval, data=df_impute, FUN=mean)

library(lattice)
xyplot(steps ~ interval | factor(wk),
       layout = c(1, 2),
       xlab="Interval",
       ylab="Number of steps",
       type="l",
       lty=1,
       data=wk_df)
```

